### 2024-03-05 paper.tex
Very quick cut of non-relevant text from sgkit paper draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please note that whilst this template provides a
% preview of the typeset manuscript for submission, it
% will not necessarily be the final publication layout.
%
% letterpaper/a4paper: US/UK paper size toggle
% num-refs/alpha-refs: numeric/author-year citation and bibliography toggle

%\documentclass[letterpaper]{oup-contemporary}
\documentclass[a4paper,num-refs]{oup-contemporary}

%%% Journal toggle; only specific options recognised.
%%% (Only "gigascience" and "general" are implemented now. Support for other journals is planned.)
\journal{gigascience}

\usepackage{graphicx}
\usepackage{siunitx}

% Additional packages used for code formatting
\usepackage{minted}
\usepackage{tcolorbox}
\usepackage{etoolbox}

% Additional packages for data model figure
\usepackage{forest}
\usepackage{tikz}
\usetikzlibrary{quotes,arrows.meta,3d}

% Local definitions to systematise tool naming
\newcommand{\sgapi}[1]{\texttt{#1}}
\newcommand{\toolname}[1]{\texttt{#1}}
\newcommand{\sgkit}{\texttt{sgkit}}

%%% Flushend: You can add this package to automatically balance the final page, but if things go awry (e.g. section contents appearing out-of-order or entire blocks or paragraphs are coloured), remove it!
% \usepackage{flushend}

% Macros etc for data model figure
\input{diagrams/data-model-preamble.tex}

% tentative
\title{Columnar VCF using Zarr}

%%% Use the \authfn to add symbols for additional footnotes, if any. 1 is reserved for correspondence emails; then continuing with 2 etc for contributions.
\author[1,\authfn{1},\authfn{2}]{First Author}
\author[2,\authfn{1},\authfn{2}]{Second Author}
\author[2]{Third Author}
\author[2,\authfn{1}]{Fourth Author}

\affil[1]{First Institution}
\affil[2]{Second Institution}

%%% Author Notes
\authnote{\authfn{1}abc@uni.edu; alphbeta@lab.edu}
\authnote{\authfn{2}Contributed equally.}

%%% Paper category
\papercat{Paper}

%%% "Short" author for running page header
\runningauthor{First et al.}

%%% Should only be set by an editor
\jvolume{00}
\jnumber{0}
\jyear{2024}

\begin{document}

\begin{frontmatter}
\maketitle

% The Abstract (250 words maximum) should be structured to
% include the following details:
% \textbf{Background}, the context and purpose of
% the study;
% \textbf{Results}, the main findings;
% \textbf{Conclusions}, brief
% summary and potential implications. Please minimize the use of abbreviations
% and do not cite references in the abstract.
% The Abstract (250 words maximum) should be structured to
% include the following details:
% \textbf{Background}, the context and purpose of % the study;
% \textbf{Results}, the main findings;
% \textbf{Conclusions}, brief summary and potential implications.

%% NOTE: this is much too long currently, but keeping for now so we
% can see which stuff migrates to the intro
\begin{abstract}
\textbf{Background:}
The largest dataset available today, the UK Biobank,
has a genotype matrix of approximately 1,000,000,000 rows and 500,000 columns,
along with metadata and quality control information of similar dimension.
Processing such matrices on a single computer is not feasible,
and computations must be distributed to complete in a reasonable timeframe.
Classical methods in genomics are inherently single-computer focused,
oriented around either streaming data through Unix pipelines or
storing the entire genotype matrix in memory.
The standard file format for genetic variation data (VCF: Variant Call Format)
is also fundamentally unsuited to distributed computation and is a severe
bottleneck at the Biobank scale.

\textbf{Results:}
key part of this approach is to use Zarr, a cloud-native
format for storing multi-dimensional numerical data, widely used in the
across the sciences to store petabyte-scale datasets, and with
multiple implementations.  We show that data stored in this format has
comparable compression levels to specialised methods to compress the
genotype matrix. We show that performing calculations on the genotype
matrix on a single computer is much more efficient at scale using sgkit than standard
approaches based on VCF, and is comparable with state-of-the-art methods
using compiled languages.

\textbf{Conclusions:}
Such a standard, based on the language-agnostic Zarr format, would enable a
diverse ecosystem of tools to efficiently access subsets of the data without
copying, and substantially reduce research costs.
\end{abstract}

\begin{keywords}
Keyword1; keyword 2; keyword 3 (Three to ten keywords representing the main content of the article)
\end{keywords}
\end{frontmatter}

%%% Key points will be printed at top of second page
\begin{keypoints*}
\begin{itemize}
\item This is the first point
\item This is the second point
\item One last point.
\end{itemize}
\end{keypoints*}

\section{Background}

\section{Results}

\subsection{Overview of narrative for draft}

\paragraph{Storing genetic variation data}
High level discussion of
strategies used to store genotype matrix and accompanying metadata.
Emphasis on access by variant, not sample. Introduce Zarr, and
refer to Fig~\ref{fig-data-storage}A to explain columnar storage,
and Fig~\ref{fig-data-storage}B to show that is competitive
with highly specialised approaches.
Note that the FC simulated dataset is probably a best-case
scenario for methods specialised to patterns in the genotype matrix.
Discuss storing additional data, and advantages of the
columnar approach (refs to genomics tools that do the same).

\paragraph{Computing on the genotype matrix}
Compressing data isn't ultimately the point - we want to use it.
We discuss the problem of running computations on the entire
genotype matrix, and how different formats limit what can be done.
We illustrate the discussion by the afdist operation from
the bcftools plugins package. This is perhaps a contrived example, but
we chose it because it requires examining the entire genotype matrix
and there is an efficient implementation available in bcftools.
It is also a good example because the standard approach of
divvying up the genome into chunks via HPC schedulers or workflow
managers is quite tricky.
We discuss the software development aspect, and how we developed
the afdist-savvy program. We discuss the inherent limitations
of the decode-to-vcf strategy, and how this inevitably
leads to bottlenecks. We discuss parallelism, and why
bcftools and genozip can make limited use of available cores.
Also that single-machine parallelism is inevitably constrained
by IO (which is why savvy's thread utilisation drops off).
Figure: \ref{fig-whole-matrix-compute}. Also
supplementary figure ~\ref{fig-whole-matrix-compute-supplemental}.

\paragraph{Subsetting the genotype matrix}
Existing work overwhelmingly emphasises accessibility by variant.
We demonstrate that sgkit/zarr is good for getting subsets
in terms of both variants and samples via afdist on various subsets.
Discuss why storing AF values for crude population groupings is
intrinsically bad, and that it should be easy to do calculations
on aribitrary subsets. Also participants regularly drop-out of
large studies like UKB - do you update the AF values each time?

\subsection{Storing genetic variation data}
Storage is the foundation on which any large-scale data-driven computation
rests: how we encode and organise data on persistent storage (the format)
fundamentally affects and limits the efficiencies of downstream calculations.
Different formats are optimised for different purposes, and there are
often many competing requirements when designing a data storage format.
% TODO is there some literature we could cite on this? I'm making up the
% term "computational format" here, presumably there is lots  prior thought
% on these ideas?
For our  purposes here, it is important to distinguish between \emph{archival}
and \emph{computational} formats. An archival format is optimised for
maximum interoperability and future-proofing, and uses the simplest and
most robust technologies to ensure the long-term usability of the data.
Reasonable compression levels are important because
of storage and transfer costs, but the efficiency of data access is a
secondary concern.
A computational format, on the other hand, is optimised
for calculations, for \emph{using} the data, today. A computational format
is designed with particular data access patterns in mind, ultimately
to make a broad range of queries and downstream computations as efficient as
possible.

% Note: a quick explanation of VCF for CS people coming in fresh
VCF~\citep{danecek2011variant} is a text-based format in which each line
describes observations
and metadata for a single variant (a position on the genome at which
variation occurs in the sample). The first eight tab-delimited
columns are mandatory, and subsequent columns describe the per-sample
genotype calls and quality control data.

[WIP: currently working on weaving together the stuff below into
some sort of narrative]

As has been noted many times~\citep[e.g.][]{kelleher2013processing,layer2016efficient,li2016bgt},

The shortcomings of VCF (and to a lesser degree, BCF) as a storage
format and the basis of efficient computation have long been
recognised
and many alternatives
have been proposed.

Purely archival formats try to get the maximum possible compression levels,
without much attention to query performance. Typically these use
highly specialised approaches to maximise compression performance
on genotype data.

These include:
TGC~\citep{deorowicz2013genome}
GTShark~\citep{deorowicz2019gtshark},
VCFShark~\citep{deorowicz2021vcfshark}.

SpVCF~\citep{lin2020sparse} is an extension of the VCF text
format that maximises compatibility, while gaining substantial
improvements in file sizes.

Many tools focus on balancing compression performance with
being able to run queries, typically outputting a VCF.
e.g.
BGT~\citep{li2016bgt},
GQT~\citep{layer2016efficient},
SAVVY~\citep{lefaive2021sparse},
XSI~\citep{wertenbroek2022xsi},
GTC~\citep{danek2018gtc},
GTRAC~\citep{tatwawadi2016gtrac}
genozip~\citep{lan2020genozip,lan2021genozip}
GBC~\citep{zhang2023gbc}.

SeqArray~\citep{zheng2017seqarray} and the GDS format~\citep{zheng2012high}
provide efficient storage and programmatic access in R.

Many recent tools are based on the Positional Burrows--Wheeler
Transform (PBWT)~\citep{durbin2014efficient}, which takes advantage
of the underlying haplotype structure of genomic data to provide
concise storage, and efficient haplotype matching.

The approach taken by \toolname{plink}~\citep{purcell2007plink,chang2015second} is to
store data in an uncompressed binary format. This has some major
advantages.

BGEN \citep{band2018bgen} is optimised for storing large-scale
genotype array data and stores probabilities, and is used
in particular by UK Biobank~\citep{bycroft2018genome}
[Should consider the "hard-call" nature of most of these formats]

Many methods working with large scale genetic
data now use the PBWT or variants of
it~\citep[e.g.][]{li2016bgt,lefaive2021sparse,wertenbroek2022xsi}. PBWTs have
disadvantages, however: data must be phased, and compression level
is quite sensitive to the presence of errors [citation? people have
worked around this]

Several tools provide querying interfaces in their CLIs
(reminiscent of SQL) to provide easy access to particular samples, or
variants~\cite{li2016bgt,layer2016efficient}.

More recently XSI~\citep{wertenbroek2022xsi} is an file format that
uses a combination of strategies to achieve excellent compression
performance, and can support some calculations directly on the
compressed format. It provides a command line interface and a
C API.

The SAV format is an extension of BCF that takes advantage
of the sparseness of large scale genomic datasets~\citep{lefaive2021sparse}.
They provide a command line interface and a C++ API.

\begin{figure}
    \begin{tabular}{c}
\resizebox{7cm}{!}{\input{diagrams/data-model.tex}} \\
\includegraphics[width=7cm]{figures/data-scaling}
\end{tabular}
\caption{Chunked compressed storage of the genotype matrix using Zarr.
(A) Schematic of the chunked array storage used in sgkit
(B) Compression performance compared with BCF, savvy and genozip
based on simulated data (see text for details).
\label{fig-data-storage}}
\end{figure}

High level discussion of
strategies used to store genotype matrix and accompanying metadata.
Emphasis on access by variant, not sample. Introduce Zarr, and
refer to Fig~\ref{fig-data-storage}A to explain columnar storage,
and Fig~\ref{fig-data-storage}B to show that is competitive
with highly specialised approaches.
Note that the FC simulated dataset is probably a best-case
scenario for methods specialised to patterns in the genotype matrix.
Discuss storing additional data, and advantages of the
columnar approach (refs to genomics tools that do the same).

\subsection{Computing on the genotype matrix}

Compressing data isn't ultimately the point - we want to use it.
We discuss the problem of running computations on the entire
genotype matrix, and how different formats limit what can be done.
We illustrate the discussion by the afdist operation from
the bcftools plugins package. This is perhaps a contrived example, but
we chose it because it requires examining the entire genotype matrix
and there is an efficient implementation available in bcftools.
It is also a good example because the standard approach of
divvying up the genome into chunks via HPC schedulers or workflow
managers is quite tricky.
We discuss the software development aspect, and how we developed
the afdist-savvy program. We discuss the inherent limitations
of the decode-to-vcf strategy, and how this inevitably
leads to bottlenecks. We discuss parallelism, and why
bcftools and genozip can make limited use of available cores.
Also that single-machine parallelism is inevitably constrained
by IO (which is why savvy's thread utilisation drops off).
Figure: \ref{fig-whole-matrix-compute}. Also
supplementary figure ~\ref{fig-whole-matrix-compute-supplemental}.

\begin{figure*}
\begin{tabular}{cc}
\includegraphics[width=6cm]{figures/whole-matrix-compute} &
\begin{tcolorbox}[width=7cm]
\begin{minted}[fontsize=\footnotesize]{python}
def sgkit_afdist(ds, num_bins=10):
    ds_vs = sgkit.variant_stats(ds).compute()
    bins = np.linspace(0, 1.0, num_bins + 1)
    bins[-1] += 0.01
    af = ds_vs.variant_allele_frequency.values[:, 1]
    pRA = 2 * af * (1 - af)
    pAA = af * af
    hets = ds_vs.variant_n_het.values
    homs = ds_vs.variant_n_hom_alt.values
    a = np.bincount(
        np.digitize(pRA, bins),
        weights=hets,
        minlength=num_bins + 1)
    b = np.bincount(
        np.digitize(pAA, bins),
        weights=homs,
        minlength=num_bins + 1)
    count = (a + b).astype(int)
    return pd.DataFrame({
        "start": bins[:-1],
        "stop": bins[1:],
        "prob_dist":
        count[1:]})

ds = sgkit.load_dataset(ds_path)
df = sgkit_afdist(ds)
\end{minted}
\end{tcolorbox}
\end{tabular}
\caption{Whole-matrix compute performance with increasing sample size.
(A) Total CPU time required to run \texttt{bcftools +afdist}
and equivalent operations in a single thread for various tools.
Elapsed time is also reported (dotted line).
(B) Relative speedup with 8 threads.
(C) The source code used for the sgkit version of afdist used in these
benchmarks. See Figure~\ref{fig-whole-matrix-compute-supplemental} for
more detailed comparisons with different numbers of threads, and comparing
the effects of HDD vs SSD storage.
\label{fig-whole-matrix-compute}}
\end{figure*}

\subsection{Subsetting the genotype matrix}
Existing work overwhelmingly emphasises accessibility by variant.
We demonstrate that sgkit/zarr is good for getting subsets
in terms of both variants and samples via afdist on various subsets.
Discuss why storing AF values for crude population groupings is
intrinsically bad, and that it should be easy to do calculations
on aribitrary subsets. Also participants regularly drop-out of
large studies like UKB - do you update the AF values each time?

\begin{figure*}
\begin{tabular}{cc}
\includegraphics[width=6cm]{figures/subset-matrix-compute} &
\begin{tcolorbox}[width=7cm]
\begin{minted}[fontsize=\footnotesize]{python}
# TODO not sure how best to illustrate this. Trying out various
# options

ds = sgkit.load_dataset(ds_path)
dss = ds.isel(
    variants=slice(...),
    samples=(slice(...)))
df = sgkit_afdist(dss)

# bcftools
bcftools view PATH -S ... -r...
| bcftools +fill-tags | bcftools +af-dist

# genozip
genocat PATH -r ... -s ...
| bcftools +fill-tags | bcftools +af-dist

# Complex query
df = pd.read_csv("sample_info.csv",
    sep="\t").set_index("sample_id")
dss = ds.isel(
    samples=df.decade[ds.sample_id.values] >= 1950)
sgkit_afdist(dss)

# Corresponding shell
% tail -n+2 sample_info.csv | awk '$2 >= 1950'
    | cut -f1 > samples.txt
bcftools view PATH -s samples.txt
    | bcftools +fill-tags | bcftools +af-dist
\end{minted}
\end{tcolorbox}
\end{tabular}
\caption{Compute performance on subsets of the matrix.
(A) Total CPU time required to run the afdist calculation for
a subset of 10 samples and 10000 variants from the middle of the matrix.
(B) Same for n/2 samples. Genozip did not run in this case for
$n > 10^4$ samples because it does not support a file to specify
sample IDs, and the command line was therefore too long for the shell
to execute.
(C) A simplified version of the source code used for these benchmarks.
\label{fig-subset-matrix-compute}}
\end{figure*}

\subsection{Distributed computation}
Distributed computation has been the reality in the analysis
of genetic variation for many years,
where data is most often provided as per-chromosome VCFs.
This chunking provides a fairly straightforward way to parallelise
computations, using HPC schedulers to split work per-file.
When more fine-grained parallelism is required, users must
provide exact genome coordinates of the genome regions
as part of their submission scripts. Combining results
is a manual operation.
Workflow engines such as
Snakemake~\cite{koster2012snakemake,molder2021sustainable}
and [TODO mention WDL, Cromwell, etc?] make such analyses
much easier, but there is still a great deal of room for
mistakes.
Per-chromosome VCFs are not practical for biobank scale
projects, however, and variant call data tends to be split
into many (still large) files.
For example, the VCFs for 150K UKB WGS data~\cite{halldorsson2022sequences}
are provided in 50kb chunks~\cite{browning2023statistical}, resulting in
hundreds of files per chromosome.
% This is horrible writing, but the point is important.
While these files provide a natural
unit for distributing work, the details of how they
are split differ and essentially there is
no interoperability among large-scale datasets because the VCFs are
so cumbersome that a separate orchestration layer is required to
access the data in parallel.

Despite this long history of de-facto distributed computation
and the clear need for better methods of accessing genetic
variation data in a distributed fashion, there is almost
no scientific literature on the topic (in stark contrast to the
study of methods to compress a VCF into a single file,
as discussed in section XXX).

[This is a really rough outline, also not sure how exactly
we sequence the ideas here. Just getting paragraph fragments
on paper. We do need to explain something of the hadoop ecosystem
to readers who are not familiar, though.]
Distributed analytics for large-scale data is a mature
techology, with many different approaches. Hadoop~\citep{white2012hadoop}
is the original method, and remains the basis of the
dominant ecosystem. HDFS. Spark [CITATION?] lets us
do more complex analytics. Parquet [CITATION?]

Although the Hadoop ecosystem has been hugely successful
in many domains, it has had a limited impact in genomics.
[List some things that have been tried for e.g variant calling
, like ADAM~\citep{nothaft2015rethinking}].
[Briefly mention two approaches using Hadoop clusters
~\citep{boufea2017managing,fan2020variant}, to not very
exiting results].

[Hadoop/Spark/Parquet are fundamentally unsuited to genetic
variation data [WHY]?]

Discuss Hail. It's awesome, has enabled current
generation of mega-scale projects. Drawbacks are the
architecture has emphasised performance etc over adaptability
and simplicity. Uses bits of Spark for X, but custom YZ .
File format is undocumented, etc.

Discuss why the sgkit architecture is suited to distributed
computation. Discuss Zarr's use in cloud stores, and existing
deployments. Discuss Dask.
We do not attempt a head-to-head performance comparison with
Hail here, but defer to Liangde's thesis~\citep{li2022efficient}
with brief summary.
Forward ref to following sections on applications to demonstate
sgkit in a distributed context.

\section{Discussion}

Outline of a possible narrative by paragraph:
\begin{itemize}
\item When VCF was introduced as part of the work of the 1000 Genomes
project the software landscape was a real mess. It has entirely
succeeded in standardising the output of variant callers, and methods
have interoperated well for the past decade. As a method of interchange
following classical file-oriented paradigms it works well enough.
It is also a good archival format: when compressed with a well established
method like gzip and combined with the well-documented description of
the text, we can be sure that files will still be readable decades
from now.

\item VCF is holding back large-scale genomics, however, and in particular the
analysis of population scale genetic variation datasets like UKB and GeL.
Describe the current state of affairs. Variant datasets are chunked up
into reasonable sized bits (10s of gigabytes) so that they can be processed
in parallel. Doing anything with them requires interacting with workflow
managers. In practise, it is not possible to directly write programs to
work across multiple large-scale datasets: although the chunks of data
are given in VCF, orchestrating how the chunks are processed and
the results combined involves substantial work, involving the workflow
manager, and the details of how chromosomes are split into chunks.
Cite examples like the phasing pipelines for UKB from Brownings etc.

\item Distributed computing is the reality, and increasingly datasets
are being mandatorily stored in public clouds as part of
Trusted Research Environments. Cloud object stores are the reality
because it is much cheaper. POSIX file systems are emulated on top of
this. Benefits of breaking up data into manageable pieces are that
compute resources can be localised to the data. Compute resources are
most effectively obtained as short-lived, small VMs that operate
as part of a large cluster. Large vertically scaled ``pet'' servers
are much more expensive, and just not feasible for the current
scale of data. Lots of other architectural
benefits from the new reality of cloud-based storage.

\item Classical methods based around Unix pipelines and streaming
files between processes on a single machine are therefore a
significant mismatch to the reality of large-scale genetic variation
datasets. The majority of methods suggested to make working
with VCF data more efficient are based on these ideas: they all
create a single large file and regard distribution of work
in a cluster as extrinsic. Most output VCF as the result of queries,
which also assumes that the VCF data can be efficiently processed.
This is a reasonable assumption for thousands of samples, but not
for larger datasets.

\item The key issue here is that VCF is not a \emph{computational}
format. Contrast VCF with plink, which stores data on disk in a
form that can be computed on directly. Plink is limited in what
can be descibed though, does not compress, and is not designed
to distributed.

\item The large-scale analysis of genetic variation data does not
need a more efficient file format, which allows files to be downloaded
somewhat more quickly, and subsets extracted somewhat more efficiently
as VCF. It needs a standard storage platform that allows data to be
processed directly without conversion, subsetted efficiently across
both the variant and sample dimensions, with first-class support
for all the necessary additional information (GT values etc),
and that is suited to the realities of today's cloud-based
storage and processing facilities.

\item We suggest that the model used by Zarr is a good choice for this.
We have shown that chunks of the genotype matrix compressed with standard
methods are competitive with highly specialised methods for
genotype data. In reality most of the storage is for non-genotype
data anyway, where everyone is essentially doing the same thing.
Columnar storage allows getting just (e.g.) the variant positions
without reading everything. Another benefit is that by storing
fields separately in different cloud objects, they can have
different permissions. So, users can be given differential access
to read parts of the dataset, and this is managed directly
by the cloud storage. Give example.

\item A major strong point of Zarr is its simplicity. All it's doing
is compressing n-dimensional chunks of numerical data and storing
them using fixed address keys. This key-value approach means that storage
is very flexible and implementation is straightforward. There
are several implementations of the Zarr protocol, and while some of these
are rudimentary, they could be improved or replaced with relatively
little effort. Zarr has been successfully used to store petabytes
of data across different sciences --- genetic variation data is largely
arrays of numerical data, and not really that different ultimately.
Zarr is cloud native.
Zarr is an open specification, built on standard components (contrast
with Hail).

\item Zarr is not perfect. Discuss limitations around ragged arrays,
etc. Forthcoming improvements in zarr v3. The whole genotype matrix
paradigm is challenged by graph genomes as well. While this is
being worked out, there is still plenty of data to analyse as a
genotype matrix.

\item Vision. An ecosystem of tools that interoperate
based on an open, cloud-based format where distribution is naturally
managed by the stored chunks. Tools can be run on different datasets
with minimal tuning, not by writing a full layer of orchestration.
Tools read the data they need, not the entire dataset. Users read
directly from a single canonical dataset, and do not maintain
their own filtered copies. We suggest that the Zarr VCF approach
we have provided would be a good starting point for designing
such a compute platform.

\end{itemize}

\section{Methods}

\subsection{Benchmarking}
Full details of the benchmarking done in
Figure~\ref{fig-data-storage} and
Figure~\ref{fig-whole-matrix-compute}.
Quick notes:

\begin{itemize}
\item Dataset is based on subsets of French-Canadian sims. We simplify
for increasingly large subsets, keeping only variant sites. We then
convert to VCF and encode to BCF using
\texttt{tskit vcf | bcftools view -O b}.
\item We used a chunk size of XX for sgkit, after some experimentation.
This gave n files of around x MB each for the ``call\_genotypes`` array.
We used the defaults for savvy and genozip.
\item For the afdist CPU time we measure the sum of the total user and
system times required to execute the full command, as reported by GNU
time. Each tool was instructed to use one thread, where the options
were provided. For genozip, the time required to compute
\texttt{genocat file.genozip | bcftools +af-dist}. Because commands
in the pipeline execute concurrently on mutiple cores, the total CPU time is
greater than the elapsed time.
\item For the threading speedup plot, we report the time required to compute
afdist with 8 threads relative to 1 thread (a speedup of 8 is therefore
the maximum). For savvy, our C++ implementation of afdist calculates
on 8 chunks of genome, and aggegates once they are completed.
Bcftools has a ``--threads`` option, but it only applies to decoding.
\item Where possible in pipelines we use uncompressed BCF
 output \texttt{-Ou} to make processing more efficient (skip printf
and parsing costs, which are substantial).
\item We do not use BCF output in genozip because it doesn't support
it directly, only VCF (supports BCF by pipeing through bcftools).
\end{itemize}

\section{Availability of source code and requirements}

\section{Data availability}

\section{Declarations}

\subsection{List of abbreviations}
% If abbreviations are used in the text they should be defined in the text at
% first use, and a list of abbreviations should be provided in alphabetical
% order.

VCF: Variant Call Format;

\subsection{Competing Interests}

\subsection{Funding}

\subsection{Author's Contributions}

\section{Acknowledgements}

\bibliography{paper}

\clearpage
\renewcommand\thefigure{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand\thetable{S\arabic{table}}
\setcounter{table}{0}

\section*{Supplementary Material}

\begin{figure*}[ht]
	\begin{center}
		\includegraphics[width=\textwidth]{figures/whole-matrix-compute-supplemental}
	\end{center}
	\caption{\label{fig-whole-matrix-compute-supplemental}
    Limitations of vertical scaling on whole-matrix afdist. The elapsed
wall-clock time required to compute afdist on the whole genotype matrix using
different numbers of threads and storage backends (top row), and the relative
speedup obtained from threads as a fraction of optimum speedup (bottom row).
[TODO more detail and link to fig 1. Also add the times for bcftools and
genozip in col 1?]
}
\end{figure*}

\clearpage
\section*{Text graveyard}

Here's a bunch of text that contains bits that might be useful and readapted.

\subsection{Chunked storage and distributed computation}
How we store genetic data fundamentally affects how computation
can be structured, and ultimately the scalability of the calculations
we wish to perform. Classically, genetic variation data is stored via one
of two strategies. The first strategy, epitomised by
VCF~\citep{danecek2011variant} encodes all the data for a given
variant in a record, and then stores these records consecutively on file
(optionally compressed). This record-oriented approach maps well to the
powerful Unix pipelines abstraction and
has proven to be very successful.
The second classical strategy, epitomised
by the \texttt{plink} BED format~\citep{purcell2007plink}, is to
store genotype data in a structured binary format that can be
accessed directly. This memory-map approach tends to
be much faster to access, but leads to larger files and is more
limited in what can be represented. There are advantages and
disadvantages to both approaches, but one problem fundamentally
limits their utility for modern datasets: neither is well suited to
scaling-out across multiple machines.

[Probably need a paragraph here on what we've done classically
by splitting by variant. This was the dominant dimension, so that
worked fine. Now we need to also think about access-by-sample]

Given the scale of modern datasets, the ability to perform
calculations in parallel on multiple CPU cores
is now a basic prerequisite, and
running concurrently on multiple computers in a cluster
is increasingly becoming necessary. [For interactive analysis? People have
been doing concurrent computing with HPC clusters forever.]
Record-oriented variant storage such as VCF is not suitable
for  distributed computing because although it is relatively
straightforward to break workloads up by variant, we cannot efficiently
split by sample (see the section on Columnar Data below for details).
Storing data in a memory-mapped manner, where the genotypes
for each variant are written consecutive consecutively to a file,
suffers from the same problem. Although it is useful to conceptualise
the genotype matrix as two-dimensional, it is important to remember
that computer memory and file storage is ultimately one-dimensonal.
For example, plink's memory-mapped data layout linearises the 2D
genotype matrix by writing [TODO check] $n$ bytes representing
the genotypes for all samples at a variant position consecutively.
Because the genotypes for a given variant index can be found
in contiguous blocks at easily computable offsets from
this initial location in memory,
accessing the genotypes for a given variant is very efficient.
However, accessing the genotypes for a given sample at every
variant position is much less efficient because each of the $m$ bytes we
must read are $n$ bytes apart in the file.
In the distributed context, this often means
that each compute node  in a
cluster must effectively read the entire file in order to obtain the
data for a subset of samples. In practise, this means that calculations
are limited by I/0 bandwidth.

A simple and elegant solution to providing efficient access to such
large matrices across multiple dimensions is to store rectangular
\emph{chunks} rather than entire rows or columns. This is illustrated
in Fig~\ref{fig-data-storage}, where we show the basic approach
used by sgkit. By storing data in these regular chunks, we
can efficiently access data for both specific variants \emph{and}
samples.

Additionally, we can compress the individual chunks [more details].

This general strategy for storing multidimensional numerical data
has been in use for many years~\citep{folk2011overview}

This maps well to distributed computing because Zarr provides a
very flexible approach to storing the chunks. By using a native cloud store
for these chunks, compute nodes can directly access the chunks that
are needed, and not read irrelevant data.
[TODO flesh out and more stuff.]

\subsubsection{Array oriented computing}

NumPy~\citep{harris2020array} has been transformational, and provides
the basic framework for scientific computing in Python.

[ Things to cite: pandas~\citep{mckinney2010data},
BioNumpy~\citep{rand2022bionumpy}, OME-Zarr~\citep{moore2023ome}]

[NOTE: we may also want a section discussing the n-dimensional array model,
but I think this section is the right place to discuss that?]

\subsubsection{Columnar binary data}

Perhaps the most important factor limiting scalability in contemporary genomics
is the continued reliance on row-wise storage of data---most often
encoded as blockwise-compressed text.
[Explain why row-wise is bad without reference to specific genomics applications.]

The Variant Call Format (VCF) is the standard method for interchanging
genomics data~\citep{danecek2011variant}.
In VCF each row describes the observations for a set of samples
at a given position along the genome, and consists of fixed columns such as the
genome position as well as per-sample columns. [More info]
VCF was introduced as part of the 1000 Genomes project, and works reasonably
well at the scale of 1000s of samples. However, when we have hundreds of
thousands of samples it becomes extremely unwieldy
. For example, the
phased genotypes for 200K samples in UK Biobank produced by Browning [cite]
require X GB of space.
(Note that this file contains no extra INFO keys; in practice a great deal
more information is stored in the VCF leading to files that are orders of
magnitude larger, and requiring them to be split into chunks, further
increasing their unwieldiness.)
Running [a simple query that requires only looking at
one column] on this file like X required Y minutes.

[TODO be more concrete here; what is the thing we're summarising?]
This slow response time to a simple query is primarily caused by the row-wise
nature of VCF (and BCF), because in order to read a single piece of
information about a particular variant site we must read \emph{all}
the information
about that site. Then, if we wish to summarise that information over
all sites, then we must read the entire dataset.

A lesser but still significant cause of inefficiency is the use of
text to store the data. [More info]

These issues are of course well known, and there are numerous projects
that seek to alleviate them. [Discuss] They're all quite bespoke to genomics data
through.

In sgkit we use Zarr which although initially introduced to store
genomics data, has now been adopted widely in [X, Y and Z].
Zarr is [description] and has [advantages]

Also to discuss (not sure how these fit in to this specific sectio
on background technologies vs existing genomics infrastructure; can
sort out later):

\begin{itemize}
\item columnar (with small records) good for I/O elimination + sequential access.
Classical locality of reference stuff as well.
\item What does this mean with ND arrays? Slippery concept. Discuss.
\item Compression improvements from columnar.
\item Poor performance of text-based VCF has been noted many times,
and lots of prior work in this space e.g. \citep{kelleher2013processing}
\item Honourable mentions for bcftools~\citep{danecek2021twelve},
htslib~\citep{bonfield2021htslib}, CRAM~\citep{bonfield2014scramble,bonfield2022cram}
which uses a columnar approach
\end{itemize}

\subsection{Storing genetic variation data}

It is important to note that the example in Fig.~\ref{fig-data-storage}
is something of a best-base
scenario in terms genomic dataset storage, in that we are only
storing genotype data (per sample, per variant),
and, because it is error-free, is very
highly compressible. Noisier genotypes will not compress quite
so well, and storing less compressible data (such as genotype
quality values) results in a large increase in the overall
storage size. This is essentially unavoidable, but the Zarr-based
approach still has some significant advantages over VCF/BCF
[basically we can store separately and come up with distinct
compression strategies based on the data type. Dunno if we
want to get into this here. This is a complex topic with
a lot of existing work which we should review and reference,
and this is maybe not the place to do it.].

[SpVCF has some good text on high-entropy QC measures. We have to
reduce this entropy or we just can't store the data]

[Why are we reviewing all this? Well, we want to make the point
that people have been trying hard to address the problem of
how to deal with large-scale genomic data for quite a while,
and have come up with lots of specialised methods. There
are a few things to keep in mind. Our approach is to use
standard methods just on the "rectangles" of variant data
and it's already quite good. You could imagine writing specialised
compression filters for genotypes, if you wanted (maybe a
discussion item though). The key thing is the \emph{interface}.
They're all tying to the command line and to classical
sequential workflows. We have Python, and distributed compute. ]

% TEXT from HSG chapter section on compressing data

% The simplest approach to compressing such data is to organise it in a
% `variant-centric' fashion, such that all observed genotypes for a given variant
% are stored in a contiguous block. Because most variants are rare, this will
% result in long runs of identical values that compress well using standard
% methods. This is the approach taken by the Variant Call
% Format~\citep{danecek2011variant}, and its more efficient binary encoding, BCF.
% The compression levels achieved by this straightforward approach are good, and
% quite competitive with more sophisticated methods. The disadvantage
% is that many queries require full decompression of the data, which can be
% prohibitively time-consuming. The SpeedGene data format and software
% library~\citep{qiao2012handling} chooses from one of three encoding methods for
% each SNP determined by the allele frequency; \cite{sambo2014compression} extend
% this idea by identifying blocks of SNPs in LD, and adding two additional
% potential encodings utilising this information.

% An alternative to this variant-centric approach is to store genotypes in a
% `sample-centric' fashion. Here, all genotypes for a particular sample are
% stored consecutively. While this breaks the simple repetition structure of data
% stored in variant-centric form, other methods can be employed to find
% compressible structure. For example, TGC~\citep{deorowicz2013genome} compresses
% sample genotypes using a custom Lempel-Ziv style approach on runs of identity
% among samples. This explicit use of LD structure results in excellent
% compression performance (for example, 32MB for all SNPs in 1000 Genomes
% chromosome 1); unfortunately, querying a dataset requires full decompression,
% making the format unsuitable for online analysis. The GQT
% toolkit~\citep{layer2016efficient} also uses the sample-centric organisation of
% genotype data, but takes a different approach to compression. Variants are
% sorted by allele frequency (resulting in longer runs of identical values within
% samples) and then compressed using an efficient bit-vector
% representation~\citep{wu2002compressing}. The resulting file-sizes are similar to
% compressed BCF, but many queries can be performed directly on the compressed
% representation and are therefore many times faster.

% The positional Burrows-Wheeler transform~\citep{durbin2014efficient}, or PBWT,
% is another sample-centric method. Building on the success of Burrows-Wheeler
% transform applications in
% genomics~\citep{langmead2009ultrafast,li2009fast,li2009soap2}, PBWT provides
% very good compression performance and efficient algorithms for finding
% haplotype matches. The algorithm builds a representation of the data based
% on sorted haplotype prefixes in linear time, and the LD structure of the data
% ensures that the sorting orders between adjacent sites change very little
% on average. The method has been successfully applied
% to phasing~\citep{loh2016reference}, detection of IBD
% segments~\citep{naseri2017ultra}, improving the performance of the Li and
% Stephens model~\citep{lunter2016fast}, and a general query engine for genotype
% data~\citep{li2016bgt}. Recent extensions include privacy preserving
% search~\citep{shimizu2016efficient} and generalisation to the setting of graph
% genomes~\citep{novak2017graph}.

\end{document}